---
title: "Common Evidence Explorer Plugin"
date: "`r Sys.Date()`"
output:
    html_document:
        number_sections: yes
        toc: yes
vignette: >
 %\VignetteIndexEntry{Common Evidence Explorer Plugin}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{css echo=FALSE}
h1  {
  margin-top: 60px;
}
```
The common evidence explorer is a shiny application and module that can be used within other shiny applications that
use OMOP standard vocabularies.

## Launching the explorer
Using the explorer works with either a database cem connector backend or a hosted url solution.
The simplest solution is to use an API as follows. This will launch the shiny app:
```{r, eval = FALSE}
baseUrl <- "https://cem.ohdsi.org/"
launchCeExplorer(apiUrl = baseUrl)
```
Alternatively, if database credentials are known the app can launch as follows:
```{r, eval = FALSE}
connectionDetails <- DatabaseConnector::createConnectionDetails("postgres", ...)
launchCeExplorer(connectionDetails = connectionDetails, cemSchema = "cem", sourceSchema = "cem_v3_sources", )
```
Consult your organisations administrator for details on connecting in database mode.
Unless you are serving a large number of requests, this mode is not required.

## Using the explorer shiny module

This approach allows you to import the Shiny module in to your shiny application of choice.

For example:

```{r, eval = FALSE}
serverFunction <- function(input, output, session) {
  backend <- CemConnector::CemWebApiBackend$new(apiUrl = "https://cem.ohdsi.org/")
  # Define a reactive that returns a dataframe with standard ingredient concepts
  ingredientConceptInput <- shiny::reactive({ tibble::tibble(conceptId = c(21604296), includeDescendants = c(1), isExcluded = c(0)) })
  # Define a reactive that returns a dataframe with standard condtion concepts
  conditionConceptInput <- shiny::reactive({ tibble::tibble(conceptId = c(4149320), includeDescendants = c(1), isExcluded = c(0)) })
  # Define a reactive that returns an integer that for looking up higher levels in the heirarchy
  siblingLookupLevelsInput <- shiny::reactive({ 1 })

  # Call the explorer module
  ceModuleServer <- ceExplorerModule("explorer", # This ID should be unqiue and match the call to ceExplorerModuleUi
                                     backend,
                                     ingredientConceptInput = ingredientConceptInput,
                                     conditionConceptInput = conditionConceptInput,
                                     siblingLookupLevelsInput = siblingLookupLevelsInput)
}

uiFunction <- function () {
  explorerTable <- ceExplorerModuleUi("exporer") # The id for the server module and ui should be the same
  shiny::fluidPage(explorerTable)
}

shiny::shinyApp(uiFunction, serverFunction)
```
The reactive objects for ingredientConceptInput and conditionConceptInput must return data.frames that are either empty or contain the column names `conceptId, includeDescendants` and `isExcluded`.