---
title: "Deploying to RSConnect"
date: "`r Sys.Date()`"
output:
    html_document:
        number_sections: yes
        toc: yes
vignette: >
 %\VignetteIndexEntry{Using CemConnector}
 %\VignetteEngine{knitr::rmarkdown}
 %\VignetteEncoding{UTF-8}
---

# Deployment
For convenience, a folder in the `extras/cem_api` folder of this package contains the file `entrypoint.R` for deployment
of the API to Posit Connect/ShinyApps.io servers.
To connect to your server in the rstudio api follow the (posit connect instructions)[https://docs.posit.co/connect/user/connecting/]

To deploy `CemConnector` first install required packages:
```{r eval = F}
install.packages(c("rsconnect", "renv"))
remotes::install_github("OHDSI/CemConnector")
```
To deploy the api then run the following after adjusting your settings:
```{r eval = F}
rsconnect::deployApi(appName = "cem_api",
                     api = "<path_to_entrypoint>", # Download the file from github or nagivate
                     envVars = list(
                       CEM_DATABASE_SERVER = "mydbmsurl",
                       CEM_DATABASE_USER = "my_user",
                       CEM_DATABASE_PASSWORD = "mysecret",
                       CEM_DATABASE_DBMS = "mydbms",
                       CEM_DATABASE_PORT = 1234,
                       CEM_DATABASE_SCHEMA = "cem"
                       CEM_DATABASE_INFO_SCHEMA = "cem_info"
                       CEM_DATABASE_VOCAB_SCHEMA = "cem_vocab"
                     ))
```
Consult (posit connect instructions)[https://docs.posit.co/] for more information on configuration of the app.